---
title: "pH_method"
output: html_document
---

```{r}
library(googledrive)
library(tidyverse)
library(ggplot2)
```

```{r}
dg_output = drive_get("Lila_Sarif_shaking_experiment", shared_drive = 'Bogar Lab')

drive_download(dg_output, type = "csv", path = "Inputs/Lila_Sarif_shaking_experiment_downloaded-20260728.csv", overwrite = TRUE)
```
```{r}
shaking_experiment_data = read.csv("Inputs/Lila_Sarif_shaking_experiment_downloaded-20260728.csv")
```

```{r}
shaking_experiment_data
```

```{r}
per_tube = shaking_experiment_data %>%
  unite("tube", c(Sample_ID, shake_time_min), sep = "_", remove = FALSE)

per_tube
  
```

How different were the repeat measurments of each prepared tube with a given method ("trial_type")?
```{r}
ggplot(per_tube, aes(x = tube, y = pH, color = trial_type, shape = repeated_measurement)) +
  geom_point()
```

```{r}

w_diff = per_tube %>%
  pivot_wider(
    names_from = repeated_measurement, # Column providing 'A' and 'B'
    values_from = !c(tube, Sample_ID, shake_time_min, trial_type)            # all but these, which are constant for a given tube
  ) %>%
  mutate(diff = abs(pH_A - pH_B))

w_diff

```
```{r}
# for time points where we were comparing methods, compare the difference between two measurements of an identical pH tube based on method
ggplot(w_diff %>% filter (shake_time_min == 20 | shake_time_min == 40), aes(x = trial_type, y = diff)) + geom_boxplot()
```
How did measurements, and differences between repeat measurements, change over shaking time?
```{r}
# with only the method that was consistent across all time points ("decant" is the method that measured temperature), compare pH of soil from same original samples over time

ggplot((per_tube %>% filter(trial_type != "as_planned")), aes(x = shake_time_min, y = pH, color = Sample_ID, shape = repeated_measurement)) + 
  geom_point() 

```
Concerns:
- CaCl2 was impure (90% salt); blank pHs should be closer to 7 than this
- Maybe not enough rinsing of electrode at 60, 80, and 100 minutes? 
  - In 20, 40, and 120 minute cases, triangles (2nd repeat measurement) are above circles (1st repeat measurement). Solution reads less acidic on second measurement, but not by much. These time points, Sarif was present for/doing many of the measurements. In 60, 80, and 100, circles (1st repeat measurement) are above triangels (2nd repeat measurement). Solution reads more acidic on second measurement, often by much more. Are remnants of soil solution lowering the pH of the blank, and subsequent soil measurements?
- 60-120 min soltions sat on the bench for a long time after shaking/decanting. 

```{r}
ggplot(per_tube, aes(x = shake_time_min, y = pH, color = trial_type, shape = Sample_ID)) + 
  geom_point() 

```



```{r}
decant_only = shaking_experiment_data %>%
  filter(trial_type == "decant")

decant_only

ggplot(decant_only, aes(x = repeated_measurement, y = temp_C_measured, color = pH, shape = Sample_ID)) +
  geom_point() + geom_jitter(width = 0.01, height = 0.01)
```

