---
title: "sample_ID_investigator"
author: "Anna Goodman"
date: "2026-06-20"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
I'm updating this file with google drive integration on 21 July 2026!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(googledrive)

library(tidyverse)

library(diffr)
```

On June 20, I manually transcribed limited metadata (Sample_ID, Core_ID, fraction) from our field data sheets so we could track down which sample_IDs we were expecting to process in the lab. These files are those transcriptions.  
```{r}
# Get and store the metadata of the file you want in a "dribble" (drive tibble)
core_data_dribble = drive_get("core_data_transcribed_20260620", shared_drive = 'Bogar Lab')
      # just specify 
      #   - the name of the file you want ("test_spreadsheet" is a csv in my Individual Research Folder)
      #   - shared_drive = 'Bogar Lab', since these files aren't in your "My Drive", which this package defaults to searching

core_data_dribble # check to make sure it found the right file, and only found ONE file with this name. If there were multiple, I think you can put more of the path to the file on drive in there
```
```{r}
sample_data_dribble = drive_get("sample_data_transcribed_IDs_20260620", shared_drive = 'Bogar Lab')

sample_data_dribble
```

```{r}
# downloading the files into the local version of the DSC_code repository
drive_download(core_data_dribble, type = "csv", path = "Inputs/core_data_transcribed_20260620_downloaded-20260721.csv")

drive_download(sample_data_dribble, type = "csv", path = "Inputs/sample_data_transcribed_IDs_20260620_downloaded-20260721.csv")
``` 

```{r}
core_data_transcribed <- read_csv("Inputs/core_data_transcribed_20260620_downloaded-20260721.csv")
sample_data_transcribed_IDs <- read_csv("Inputs/sample_data_transcribed_IDs_20260620_downloaded-20260721.csv")
```

```{r}
# make vectors of Sample_ID from each source
core_data_sIDs = core_data_transcribed$Sample_ID
sorted_sIDs_core = sort(core_data_sIDs)

sample_data_sIDs = sample_data_transcribed_IDs$Sample_ID
sorted_sIDs_sample = sort(sample_data_sIDs)
```

## Compare vectors of Sample_IDs only
```{r}
  ## which Sample_IDs appear in one but not the other?
setdiff(sorted_sIDs_core, sorted_sIDs_sample)

# several; not necesarily an issue because deep samples shouldn't appear on sample data sheets from the field
```

```{r}
  ## do any appear multiple times in either source?
sorted_sIDs_core_no_nc = sorted_sIDs_core[!sorted_sIDs_core %in% c("not_collected")] # remove all instances of "not_collected" from Sample_ID column

length(sorted_sIDs_core_no_nc) == length(unique(sorted_sIDs_core_no_nc)) # should return TRUE if no duplicates
length(sorted_sIDs_sample) == length(unique(sorted_sIDs_sample)) # should return TRUE if no duplicates

#looks like all uniqe for both

        # the same thing another way:
duplicated(sorted_sIDs_core_no_nc) # all false if no duplicates
duplicated(sorted_sIDs_sample)

# yep all good
```

```{r}
# sort whole dataframes from each source by Sample_ID
core_data_transcribed = core_data_transcribed %>% arrange(Sample_ID)

sample_data_transcribed = sample_data_transcribed_IDs %>% arrange(Sample_ID)
```

## Compare dataframes with Core_ID, fraction, site
```{r}
# compare whole dataframes
write_csv(core_data_transcribed, "Results/sorted_sIDs_core.csv")
sIDs_core_file = "Results/sorted_sIDs_core.csv"

sample_data_transcribed_for_diffr = sample_data_transcribed %>% select(-c(Scribe_initials, Subsampling_team))
write_csv(sample_data_transcribed_for_diffr, "Results/sorted_sIDs_sample.csv")
sIDs_sample_file = "Results/sorted_sIDs_sample.csv"

diffr(sIDs_core_file, sIDs_sample_file)
```
 
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .  
  .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .               
 .    
 .  
 .  
 .  
 .  
 .  
  .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .               
 .    
 .  
 .  
 .  
 .  
 .  
  .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  

- ^ Most of these differences are non-issues; core data includes samples w/ depth > 50cm, and all SOAP Os -- none of which the subsampling crews handled at any point

- Several Core_ID discrepancies, but core was not reliably communicated to subsampling teams; trust core data Core_IDs and disregard sample data Core_IDs; match based on Sample_ID. Explored further below. 

- Only real issue I see is double use of DSC_041 and DSC_042
  - on original copy of NEWT core data sheet with U-01 and U-02, DSC_041 is assigned to the O fraction of U-K, then crossed out. DSC_042 is assigned to the O fraction of U-X, then crossed out. I didn't transcribe crossed-out values.  
  - on an original copy of NEWT core data sheet with U-M and U-Z, DSC_041 and DSC_042 are assigned to deep U-Z samples. (150-200cm and 200-300cm, respectively.) These were both part of segment 2 from U-Z, which does not have arrival time recorded but happened sometime between U0Z segment 1 arriving at 13:33 on 17 June 2026 and when segment 1 of U-01 arrived at 2:40PM the same day.
  - from my memory, U-K and U-X had O horizons sampled by Glade before we broke ground on U-M, during geoprobe failure. Subsampling_start_time is not recorded for DSC_041 on sample data, but it's 13:02 for DSC_042. Seems likely that these Sample_IDs were assigned to the Os but not crossed off on the master list, so the IDs were used again during U-M sampling.  

```{r}
# trim and format whole dataframes to make them comparable

  # remove Y and Z sample IDs from core data, which should not appear on field subsampling data sheets
core_data_transcribed = core_data_transcribed %>%
  filter(fraction != "50-100_cm") %>%
  filter(fraction != "100-150_cm") %>%
  filter(fraction != "150-200_cm") %>%
  filter(fraction != "200-300_cm") %>%
  filter(fraction != "300-400_cm") %>%
  filter(fraction != "400-500_cm")

  # arrange both dataframes by core, then fraction, then Sample_ID, so they align nicely in diffr
sample_data_transcribed_by_core = sample_data_transcribed_for_diffr %>% 
  arrange(Sample_ID) %>%
  arrange(fraction) %>%
  arrange(Core_ID)
core_data_transcribed_by_core = core_data_transcribed %>% 
  arrange(Sample_ID) %>%
  arrange(fraction) %>%
  arrange(Core_ID)
```

```{r}
# compare whole dataframes again, this time only with samples they should share
write_csv(core_data_transcribed_by_core, "Results/sorted_sIDs_core.csv")
sIDs_core_file = "Results/sorted_sIDs_core.csv"

write_csv(sample_data_transcribed_by_core, "Results/sorted_sIDs_sample.csv")
sIDs_sample_file = "Results/sorted_sIDs_sample.csv"

diffr(sIDs_core_file, sIDs_sample_file)
```
 
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .   .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
 .  
# Explanations for all remaining discrepancies:

### DSC_008 (B-E) missing from sample data: 
- it was an O at SOAP. Glade collected surface organic matter for each core at SOAP and gave these collections/notes Sample_IDs, but no SOAP cores had enough surface material to justify processing it as an O horizon. Subsampling teams did not see any O samples at SOAP. 
  
### DSC_084 is B-F in core data, B-R in sample data:
- Core_ID was not communicated reliably or thoroughly with subsampling teams (while Sample_ID was). Both sources describe this sample as 15-30 cm. No 15-30 cm sample for B-F is described in sample data. I conclude that DSC_084 is from B-F, and was misattributed to B-R in sample data.
  
### DSC_011 (B-F) is missing from sample data:
- it was an O at SOAP. See above. 
  
### DSC_091 is B-H in core data, B-S in sample data:
- Core_ID was not communicated reliably or thoroughly with subsampling teams (while Sample_ID was). Both sources describe this sample as 0-15 cm. No 0-15 cm sample for B-H is described in sample data. I conclude that DSC_091 is from B-H, and was misattributed to B-S in sample data.
  
### DSC_087 (B-H) is missing from sample data:
- it was an O at SOAP. See above. 

### DSC_079 (B-R) is missing from sample data:
- it was an O at SOAP. See above. 

### DSC_083 (B-S)= is missing from sample data:
- it was an O at SOAP. See above. 

### DSC_088 (B-U)= is missing from sample data:
- it was an O at SOAP. See above. 

### DSC_048 is U-J in core data, U-02 in sample data:
- Core_ID was not communicated reliably or thoroughly with subsampling teams (while Sample_ID was). Both sources describe this sample as 0-15 cm. No 0-15 cm sample for U-J is described in sample data. I conclude that DSC_048 is from U-J, and was misattributed to U-02 in sample data.

### DSC_041 and DSC_042 appear in sample data but not [this O -> 50cm only version of] core data; see notes after last diffr.

### DSC_052 and DSC_053 are U-W in core data, core not specified in sample data:
- Core_ID was not communicated reliably or thoroughly with subsampling teams (while Sample_ID was). DSC_052 is 15-30 cm in both sources and DSC_053 is 30-50 cm in both sources. I conclude that DSC_052 and DSC_053 are from U-W.  
.  

# Actions Anna will take:

### Cross out incorrect Core_IDs on sample data sheets from June field sampling and write correct ones, based on descisions described above. I'll do this in a red pen and date my corrections on the page.
- DSC_084: correcting B-R -> B-F
- DSC_091: correcting B-S -> B-H
- DSC_048: correcting U-02 -> U-J

### Write in missing Core_IDs on sample data sheets from June field sampling. I'll do this with the same red pen and date my corrections on the page. 
- DSC_052: specifying U-W
- DSC_053: specifying U-W

### Remove material and cross out all data from U-K and U-X, ensuring that [U-Z 150-200_cm] is the only DSC_041 and [U-Z 200-300_cm] is the only DSC_042.
- Search for all subsampling vessels that exist from the field processing of the U-K and U-X O horizons, and:
  - THOROUGHLY cross/black out their Sample_ID labels
  - write their Core_ID on them
  - throw them away?
- Cross out the Sample_IDs (DSC_041 and DSC_042) where they are written on June field sampling sample data sheets with my red pen. Date my corrections on the page. 
  
### Make new digital copies of all field datasheets and upload to the drive. Move original digital copies of the field datasheets to a new folder, clearly labeled as archival. 