---
title: "subsample_and_photo_inventory"
output: html_document
---

```{r}
library(googledrive)

library(tidyverse)

library(diffr)
```

This script is for 
- making a csv that inventories the subsamples and photos we have for each DSC sample, with *traceable* data origins
- automating the photo inventory


#Get a list of all the sample IDs
- Glade did this manually sometime in late June or early July, after the YZ sample processing was done. We'll use his list
```{r}
# Get and store the drive metadata of the file Glade made
Glades_sID_list = drive_get("what samples exist and what are they called", shared_drive = 'Bogar Lab')

Glades_sID_list # check to make sure we found the right file
```
```{r}
drive_download(Glades_sID_list, type = "csv", path = "Inputs/what samples exist and what are they called_downloaded-20260721.csv")
```

```{r}
sID_list_source = read.csv("Inputs/what samples exist and what are they called_downloaded-20260721.csv")
head(sID_list_source)

sID_list = sID_list_source$Sample_ID
```

# Get a list of all the photos of each type that we have?
```{r}
photo_folder_lab = drive_reveal(drive_get("LAB/PHOTOS", shared_drive = 'Bogar Lab'), "path")
```

```{r}
drive_ls(photo_folder_lab, shared_drive = 'Bogar Lab')
```

```{r}
photo_folder_lab_lrr = drive_reveal(drive_get("LAB/PHOTOS/L-ROOT_and_ROCK_photos", shared_drive = 'Bogar Lab'), "path")

lab_lrr_dribble = drive_ls(photo_folder_lab_lrr, shared_drive = 'Bogar Lab')
lab_lrr_dribble$name
```
```{r}
lab_lrr_photo_names = as.data.frame(lab_lrr_dribble$name) # store the file name column only in a new dataframe

head(lab_lrr_photo_names)

# digest file names into Sample_ID, content specifiers (photo type?), and the file extension
lab_lrr_photo_names = lab_lrr_photo_names %>% 
  rename(name = `lab_lrr_dribble$name`) %>%
  separate(name, c("filename", "file_extension"), sep = "K.") %>% # I don't know why it won't let me split on the "." :(
  separate(filename, c("DSC", "sIDnum", "content1", "content2"), sep = "_") %>%
  unite(Sample_ID, c(DSC, sIDnum), sep = "_")

head(lab_lrr_photo_names)

```
# make blank template for inventory sheet!
```{r}
template = as.data.frame(sID_list)

template = template %>%
  mutate(photo_Lroot = "not_accounted_for") %>%
  mutate(photo_ROCK = "not_accounted_for") %>%
  mutate(photo_sampleClosed = "not_accounted_for") %>%
  mutate(photo_sampleOpen = "not_accounted_for") %>%
  rename(Sample_ID = sID_list)

head(template)
```
 -------------------------
version of the lab_lrr_photo_names df creation that keeps all content in one column
```{r}
lab_lrr_photo_names = as.data.frame(lab_lrr_dribble$name) # store the file name column only in a new dataframe

head(lab_lrr_photo_names)

# digest file names into Sample_ID, content specifiers (photo type?), and the file extension
lab_lrr_photo_names = lab_lrr_photo_names %>% 
  rename(name = `lab_lrr_dribble$name`) %>%
  separate(name, c("DSC", "sIDnum", "content1", "content2"), sep = "_") %>%
  unite(Sample_ID, c(DSC, sIDnum), sep = "_") %>%
  unite(content, c(content1, content2), sep = "_")

head(lab_lrr_photo_names)

```
adds info about just lab root and rock photos to a blank inventory template
```{r}
w_lab_lrr = template %>% 
  left_join(lab_lrr_photo_names, by = "Sample_ID") %>%
  mutate(photo_Lroot = case_when(
    str_detect(content, "L-ROOT") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  mutate(photo_ROCK = case_when(
    str_detect(content, "ROCK") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   ))

View(w_lab_lrr)
```


--------------------
Can I do this but instead of lab_lrr_photo_names, build a dataframe with the name and content of all relevant photos? 






# Get a list of all the photos we have!

Get the location of all of our photo folders on google drive
```{r}
# lab photos
photo_folder_lab = drive_reveal(
  drive_get("JUNE_original_data_notes_photos/LAB/PHOTOS", shared_drive = 'Bogar Lab'), 
  "path")

# field photos
photo_folder_field = drive_reveal(
  drive_get("JUNE_original_data_notes_photos/FIELD/PHOTOS", shared_drive = 'Bogar Lab'), 
  "path")

# drive_get() searches the specified shared google drive (Bogar Lab) attached to my drive account for the file I specified
# drive_reveal() finds and stores even more specific metadata; in this case, the full file path. 
```
Get a list of all the subfolders in those photo folders
```{r}
photo_subfolders_lab = drive_ls(photo_folder_lab, shared_drive = 'Bogar Lab')
photo_subfolders_field = drive_ls(photo_folder_field, shared_drive = 'Bogar Lab')

```
bind those lists together; now we have a list of all the internal google drive identifiers of 
```{r}
photo_subfolder_IDs = append(photo_subfolders_lab$id, photo_subfolders_field$id) 
```

```{r}
photo_subfolder_contents = lapply(as_id(photo_subfolder_IDs),
  drive_ls)

photo_subfolder_contents # ok this is a list of dribbles which each contains a list of filenames

photo_filenames = c()
for (var in photo_subfolder_contents) {
  photo_filenames = append(photo_filenames, var$name)
}
photo_filenames # ok yay we have a vector of all the photo filenames!
```


version of the lab_lrr_photo_names df creation that keeps all content in one column
```{r}
photo_names_df = as.data.frame(photo_filenames) # store the file names in a dataframe

head(photo_names_df)
```

```{r}
# digest file names into Sample_ID, content specifiers (photo type?), and the file extension
photo_content_df = photo_names_df %>% 
  rename(name = photo_filenames) %>%
  separate(name, c("DSC", "sIDnum", "content1", "content2"), sep = "_") %>% # separate by _. Having multiple content columns is necesary for cases where there's an underscore in the file name after the one that separates the sample ID from the content -- but I'll just unite them again below. This also generates some _NA when there isn't an underscore in the latter part of the content
  unite(Sample_ID, c(DSC, sIDnum), sep = "_") %>% # puts DSC_[sample number] back together as the full sample id
  unite(content, c(content1, content2), sep = "_") %>% # puts everything else back together
  mutate(content = str_remove(content, "_NA")) # take out those extra NAs I generated above for readability later

head(photo_content_df)

```
Make the photo inventory!
```{r}
photo_inventory = template %>% 
  
  left_join(photo_content_df, by = "Sample_ID") %>% # Joins the photo names dataframe to our template; makes a dataframe with all the possible DSC sample IDs in one column and the rest of the photo filenames in another (the "content" column). Sample_IDs are repeated for each unique photo, and are present with "NA" in the content columns if we have no photo

  group_by(Sample_ID) %>%
  summarise(
    content = str_flatten(content, ";")) %>% # this and the line before it work together to make the dataframe have each sample ID only once, and makes the content column a long messy string of all the remaining photo name content

  # log which sample IDs have an Lroot photo
  mutate(photo_Lroot = case_when(
    str_detect(content, "L-ROOT") == TRUE ~ "in_drive",
    str_detect(content, "Lroot") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
    # log which sample IDs have a rock photo
  mutate(photo_ROCK = case_when(
    str_detect(content, "ROCK") == TRUE ~ "in_drive",
    str_detect(content, "rock") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
  # log which sample IDs have a sample-closed photo
  mutate(photo_sampleClosed = case_when(
    str_detect(content, "sample-closed") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
  # log which sample IDs have a sample-open photo
  mutate(photo_sampleOpen = case_when(
    str_detect(content, "sample-open") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   ))

View(photo_inventory)
```



