---
title: "subsample_and_photo_inventory"
output: html_document
---

```{r}
library(googledrive)

library(tidyverse)
```

This script is for 
- making a csv that inventories the subsamples and photos we have for each DSC sample, with *traceable* data origins
- automating the photo inventory


#Get a list of all the sample IDs
- Glade did this manually sometime in late June or early July, after the YZ sample processing was done. We'll use his list
```{r}
# Get and store the drive metadata of the file Glade made
Glades_sID_list = drive_get("what samples exist and what are they called", shared_drive = 'Bogar Lab')

Glades_sID_list # check to make sure we found the right file
```
```{r}
drive_download(Glades_sID_list, type = "csv", path = "Inputs/what samples exist and what are they called_downloaded-20260721.csv")
```

```{r}
sID_list_source = read.csv("Inputs/what samples exist and what are they called_downloaded-20260721.csv")
head(sID_list_source)

sID_list = sID_list_source$Sample_ID
```

# make blank template for inventory sheet!
```{r}
template = as.data.frame(sID_list)

template = template %>%
  mutate(photo_Lroot = "not_accounted_for") %>%
  mutate(photo_ROCK = "not_accounted_for") %>%
  mutate(photo_sampleClosed = "not_accounted_for") %>%
  mutate(photo_sampleOpen = "not_accounted_for") %>%
  rename(Sample_ID = sID_list)

head(template)
```


# Get a list of all the photos we have!

Get the location of all of our photo folders on google drive
```{r}
# lab photos
photo_folder_lab = drive_reveal(
  drive_get("JUNE_original_data_notes_photos/LAB/PHOTOS", shared_drive = 'Bogar Lab'), 
  "path")

# field photos
photo_folder_field = drive_reveal(
  drive_get("JUNE_original_data_notes_photos/FIELD/PHOTOS", shared_drive = 'Bogar Lab'), 
  "path")

# drive_get() searches the specified shared google drive (Bogar Lab) attached to my drive account for the file I specified
# drive_reveal() finds and stores even more specific metadata; in this case, the full file path. 
```
Get a list of all the subfolders in those photo folders
```{r}
photo_subfolders_lab = drive_ls(photo_folder_lab, shared_drive = 'Bogar Lab')
photo_subfolders_field = drive_ls(photo_folder_field, shared_drive = 'Bogar Lab')

```
bind those lists together; now we have a list of all the internal google drive identifiers of 
```{r}
photo_subfolder_IDs = append(photo_subfolders_lab$id, photo_subfolders_field$id) 
```

```{r}
photo_subfolder_contents = lapply(as_id(photo_subfolder_IDs),
  drive_ls)

photo_subfolder_contents # ok this is a list of dribbles which each contains a list of filenames

photo_filenames = c()
for (var in photo_subfolder_contents) {
  photo_filenames = append(photo_filenames, var$name)
}
photo_filenames # ok yay we have a vector of all the photo filenames!
```


version of the lab_lrr_photo_names df creation that keeps all content in one column
```{r}
photo_names_df = as.data.frame(photo_filenames) # store the file names in a dataframe

head(photo_names_df)
```

```{r}
# digest file names into Sample_ID, content specifiers (photo type?), and the file extension
photo_content_df = photo_names_df %>% 
  rename(name = photo_filenames) %>%
  separate(name, c("DSC", "sIDnum", "content1", "content2"), sep = "_") %>% # separate by _. Having multiple content columns is necesary for cases where there's an underscore in the file name after the one that separates the sample ID from the content -- but I'll just unite them again below. This also generates some _NA when there isn't an underscore in the latter part of the content
  unite(Sample_ID, c(DSC, sIDnum), sep = "_") %>% # puts DSC_[sample number] back together as the full sample id
  unite(content, c(content1, content2), sep = "_") %>% # puts everything else back together
  mutate(content = str_remove(content, "_NA")) # take out those extra NAs I generated above for readability later

head(photo_content_df)

```
Make the photo inventory!
```{r}
photo_inventory = template %>% 
  
  left_join(photo_content_df, by = "Sample_ID") %>% # Joins the photo names dataframe to our template; makes a dataframe with all the possible DSC sample IDs in one column and the rest of the photo filenames in another (the "content" column). Sample_IDs are repeated for each unique photo, and are present with "NA" in the content columns if we have no photo

  group_by(Sample_ID) %>%
  summarise(
    content = str_flatten(content, ";")) %>% # this and the line before it work together to make the dataframe have each sample ID only once, and makes the content column a long messy string of all the remaining photo name content

  # log which sample IDs have an Lroot photo
  mutate(photo_Lroot = case_when(
    str_detect(content, "L-ROOT") == TRUE ~ "in_drive",
    str_detect(content, "Lroot") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
    # log which sample IDs have a rock photo
  mutate(photo_ROCK = case_when(
    str_detect(content, "ROCK") == TRUE ~ "in_drive",
    str_detect(content, "rock") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
  # log which sample IDs have a sample-closed photo
  mutate(photo_sampleClosed = case_when(
    str_detect(content, "sample-closed") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
  # log which sample IDs have a sample-open photo
  mutate(photo_sampleOpen = case_when(
    str_detect(content, "sample-open") == TRUE ~ "in_drive",
    .default = "not_accounted_for",
   )) %>%
  
  select(!content) # remove the big messy content column

View(photo_inventory)
```

ok now that we know which photos we don't have, we should figure out who's suppoused to have them

# Add photo taker initials from transcribed sample data
```{r}
# dribble searching for transcribed sample data for x samples
X_data_path = drive_reveal(
  drive_get("X_sample_data_transcribed", shared_drive = 'Bogar Lab'), 
  "path")

# dribble searching for transcribed sample data for y samples
Y_data_path = drive_reveal(
  drive_get("YZ_sample_data_transcribed", shared_drive = 'Bogar Lab'), 
  "path")

# check that each of these has the one, correct file in it 
X_data_path
Y_data_path
#ok nice
```
download the transcribed data from google drive -- and save the path dribbles, for resolving future mix-ups?
```{r}
drive_download(X_data_path, type = "csv", path = "Inputs/X_sample_data_transcribed_downloaded-20260722.csv")
drive_download(Y_data_path, type = "csv", path = "Inputs/YZ_sample_data_transcribed_downloaded-20260722.csv")

write_csv(X_data_path, "Inputs/DRIBBLE_X_sample_data_transcribed_downloaded-20260722.csv") # should we be saving this metadata, in case file names get changed in google drive?
write_csv(Y_data_path, "Inputs/DRIBBLE_YZ_sample_data_transcribed_downloaded-20260722.csv") # should we be saving this metadata, in case file names get changed in google drive?

```

read the transcribed data into the R enviornment
```{r}
x_data = read_csv("Inputs/X_sample_data_transcribed_downloaded-20260722.csv", col_names = FALSE)
y_data = read_csv("Inputs/YZ_sample_data_transcribed_downloaded-20260722.csv", col_names = FALSE)

#Rename the columns with their Sample_IDs

sampleIDrow = which(x_data[[1]] == "Sample_ID") # find which row # contains all the Sample_IDs
colnames(x_data) = as.character(x_data[sampleIDrow,]) # make Sample_IDs the column names
x_data = x_data %>%
  rename(futureColumn = Sample_ID) %>% # rename the first column, which actually contains all the future column names (variable names? measurement names?)
  slice(-sampleIDrow) # important to take the sample_ID row out so pivoting doesn't get confused later
head(x_data)

sampleIDrow = which(y_data[[1]] == "Sample_ID")
colnames(y_data) = as.character(y_data[sampleIDrow,])
y_data = y_data %>%
  rename(futureColumn = Sample_ID) %>%
  slice(-sampleIDrow)

head(y_data)
```

Pivot to get sample_ID as its own column and all the measurements in rows
  -- long first
```{r}
x_data_long = x_data %>%
  pivot_longer(cols = !futureColumn, 
               names_to = "Sample_ID", 
               values_to = "value"
               ) 
head(x_data_long)

y_data_long = y_data %>%
  pivot_longer(cols = !futureColumn, 
               names_to = "Sample_ID", 
               values_to = "value"
               ) 
head(y_data_long)
```
  -- then wide
```{r}
x_data_wide = x_data_long %>%
  pivot_wider(names_from = futureColumn, 
              values_from = value)

head(x_data_wide)

y_data_wide = y_data_long %>%
  pivot_wider(names_from = futureColumn, 
              values_from = value)

head(y_data_wide)
```
NOW add photo takers to inventory
```{r}
y_photo_takers = y_data_wide %>%
  select(c(Sample_ID, Lroot_photo_by, rock_photo_by, pre_trim_photo_by, post_trim_photo_by))

photo_inventory_w_yinitials = photo_inventory %>%
  left_join(y_photo_takers) %>%
  mutate(sampleOpen_photo_by = post_trim_photo_by,
         sampleClosed_photo_by = pre_trim_photo_by) %>%
  select(!c(pre_trim_photo_by, post_trim_photo_by))

x_photo_takers = x_data_wide %>%
  select(c(Sample_ID, Lroot_and_rock_photos_by))

photo_inventory_w_initials = photo_inventory_w_yinitials %>%
  left_join(x_photo_takers) %>%
  mutate(
    rock_photo_by = case_when( #rock_photo_by is currently whatever it was on the YZ sample data, or (ideally) NA (the data class) for X samples
    is.na(rock_photo_by) & !is.na(Lroot_and_rock_photos_by) ~ Lroot_and_rock_photos_by, # where rock_photo_by is of class NA (not "N/A" as a character entry, as it would be if evan transcribed it that way), but the X sample data has a set of initials in Lroot_and_rock_photos_by for this sample ID, put those initials in rock_photo_by
    !is.na(rock_photo_by) & is.na(Lroot_and_rock_photos_by) ~ rock_photo_by, # where there were rock photo initials recorded on the YZ data sheets but not on the X data sheets, use the data from the YZ data sheets
    !is.na(rock_photo_by) & !is.na(Lroot_and_rock_photos_by) ~ "error: initials recorded on both X and YZ sample data",
    is.na(rock_photo_by) & is.na(Lroot_and_rock_photos_by) ~ "no photo taker recorded in sample data"
                            ),
    Lroot_photo_by = case_when( #Lroot_photo_by is currently whatever it was on the YZ sample data, or (ideally) NA (the data class) for X samples
    is.na(Lroot_photo_by) & !is.na(Lroot_and_rock_photos_by) ~ Lroot_and_rock_photos_by, # where rock_photo_by is of class NA (not "N/A" as a character entry, as it would be if evan transcribed it that way), but the X sample data has a set of initials in Lroot_and_rock_photos_by for this sample ID, put those initials in Lroot_photo_by
    !is.na(Lroot_photo_by) & is.na(Lroot_and_rock_photos_by) ~ Lroot_photo_by, # where there were rock photo initials recorded on the YZ data sheets but not on the X data sheets, use the data from the YZ data sheets
    !is.na(Lroot_photo_by) & !is.na(Lroot_and_rock_photos_by) ~ "error: initials recorded on both X and YZ sample data",
    is.na(Lroot_photo_by) & is.na(Lroot_and_rock_photos_by) ~ "no photo taker recorded in sample data"
                            )
    )%>%
  select(!Lroot_and_rock_photos_by)
  
View(photo_inventory_w_initials)
```




