---
title: "Early DSC June data exploration with plotSPC()"
output: html_document
---

```{r}
library(tidyverse)
library(aqp)
```
Read in metadata, rename columns to play nice with aqp()
```{r}
psm = read_csv("../../TidyData/Metadata/make_per_sample_metadata/Results/DSC_per_sample_metadata_compiled-20260819.csv")
colnames(psm)

psm = psm %>% 
  rename(name = Sample_ID,
         id = Core_ID,
         top = sample_top_depth,
         bottom = sample_bottom_depth) %>%
  filter(!is.na(top))

depths(psm) <- id ~ top + bottom

```
# Make crude soil moisture plots
```{r}
plotSPC(psm, 
        name = 'name',
        color = 'FM_moisture_percent' )
```

```{r}
# mask the all-root sample because it's throwing off the color scale
h = horizons(psm)
h$FM_moisture_percent = as.numeric(h$FM_moisture_percent)
h = h %>% mutate(
  FM_moisture_percent = case_when(
    FM_moisture_percent > 60 ~ NA,
    FM_moisture_percent < 60 ~ FM_moisture_percent
  ))
horizons(psm) = h


plotSPC(psm, 
        name = 'name',
        color = 'FM_moisture_percent' )
```

# Make crude pH plots

```{r}
# use average of two pH measurements for each sample
h = horizons(psm)
h = h %>% mutate(
  pH = (pH_1st_measurement + pH_2nd_measurement)/ 2
  )
horizons(psm) = h

plotSPC(psm, 
        name = 'name',
        color = 'pH', 
        col.palette = hcl.colors(14, "Harmonic"))
```
```{r}
library(aqp)
library(dplyr)

# 1. Update pH horizon values
h <- horizons(psm) %>% 
  mutate(pH = (pH_1st_measurement + pH_2nd_measurement) / 2)
horizons(psm) <- h

# 2. Extract Site and microsite from horizons (getting 1 row per profile)
h_site_info <- horizons(psm) %>%
  group_by(id) %>% # replace 'id' with your profile ID column if different (e.g., idcol(psm))
  dplyr::slice(1) %>%
  ungroup()

# 3. Build hierarchical plot order (Site, then microsite)
hierarchical_order <- order(h_site_info$Site, h_site_info$microsite)

# 4. Plot profiles grouped by Site and microsite
plotSPC(
  psm, 
  name = 'name',
  color = 'pH', 
  col.palette = hcl.colors(14, "Harmonic"),
  plot.order = hierarchical_order
)
```

# Make crude Lroot plot 
```{r}

# make Lroot mass numeric
h = horizons(psm)
h = h %>% mutate(
  Lroot_full_mass_g = as.numeric(Lroot_full_mass_g),
  Lroot_empty_mass_g = as.numeric(Lroot_empty_mass_g),
  Lroot_mass = Lroot_full_mass_g - Lroot_empty_mass_g, 
  thickness = bottom - top,
  Lroot_relative = Lroot_mass / thickness

) 
horizons(psm) = h


plotSPC(psm, 
        name = 'name',
        color = 'Lroot_relative' )

# ok some of these NAs should be 0s (where there were no roots so nothing was written) but some of them are because we forgot to weight; need to go into notes and tidy this
```
# Make crude rock mass plot
```{r}

# make rock mass numeric
h = horizons(psm)
h = h %>% mutate(
  rock_full_mass_g = as.numeric(rock_full_mass_g),
  rock_empty_mass_g = as.numeric(rock_empty_mass_g),
  rock_mass = rock_full_mass_g - rock_empty_mass_g,
  thickness = bottom - top,
  volume = thickness * 4.8 * pi) %>% # double check that ID of tubes was 4.8cm 

  filter(rock_mass >= 0) %>%
  mutate(rock_relative = rock_mass / volume
  ) %>%
  filter(rock_relative != Inf) # temporary issue with 0 thicknesses from os, bc core data not yet entered with o thicknesses

horizons(psm) <- h


plotSPC(psm, 
        name = 'name',
        color = 'rock_relative',
        col.label = 'g 4mm+ rock per cm^3')
```

ok data cleaning to-do
- make depths actual; use core data to adjust bottom depths of bottommost samples. Account for shoes
- deal with DSC_106; it needs 3 parts to be in this depth profile and they need depths assigned
- integrate core data!!
    - add O thicknesses
- clean up NAs in lroot and rock masses; which are actually 0?
- add Lroot and rock masses (with empty masses subtracted) to main psm csv
    






