---
title: "tidy_moisture"
output: html_document
---

NOT COMPLETE AS OF 14 AUG 2026. only oven-dry data tidied. air-dry data loaded, not tidied. no correction or notation for the handful of OD measurements which are unuseable; see anna's notes and the "qualcheck" tab of the google drive file "oven_dry_masses_JulyAug2026"

```{r}
library(googledrive)
library(tidyverse)
```

```{r}

name_of_transcription_file_on_drive = "oven_dry_masses_JulyAug2026" # copy the file's name exactly as it appears on google drive

dribble = drive_get(name_of_transcription_file_on_drive, shared_drive = 'Bogar Lab') # get necesary data for downloading

local_path <- paste0("Inputs/", 
                     name_of_transcription_file_on_drive, 
                     "_downloaded-", 
                     format(Sys.Date(), "%Y%m%d"), 
                     ".csv") #define what this local file name will be, with today's date

drive_download(dribble, type = "csv", path = local_path, overwrite = TRUE)

od_transcribed_data = read_csv(local_path)

```
# Oven dry moisture data
General tidy 
```{r}
od_measures = od_transcribed_data %>%
  
  #exclude rows that don't contain oven-dry mass measurements
  mutate (oven_dry_plus_tin_mass_g = as.numeric(oven_dry_plus_tin_mass_g)) %>%
  filter(!is.na(oven_dry_plus_tin_mass_g)) %>%
  
  #rename the unamed notes column
  rename(od_notes = X)

```

Calculate FM_moisture_percent and AD_Moisture_percent from oven dry sheet (small tins)
```{r}
od_moisture_tidy = od_measures %>%
  mutate(
         FM_mass_g = field_moist_plus_tin_mass_g - empty_tin_mass_g,
         AD_mass_g = air_dry_plus_tin_mass_g - empty_tin_mass_g,
         OD_mass_g = oven_dry_plus_tin_mass_g - empty_tin_mass_g,
         
         AD_Factor = AD_mass_g / FM_mass_g, 
         
         OD_Factor = OD_mass_g / FM_mass_g, 
         
         diff = AD_Factor - OD_Factor,
         
         FM_moisture_percent =
             ((FM_mass_g - OD_mass_g) / (OD_mass_g)) * 100,
         AD_moisture_percent = 
           ((AD_mass_g - OD_mass_g) / (OD_mass_g)) * 100
         )
  
```

export
```{r}

local_path_tidydata <- paste0("Results/", 
                     "od_moisture_tidied-", 
                     format(Sys.Date(), "%Y%m%d"), 
                     ".csv") #define what this local file name will be, with today's date

write_csv(od_moisture_tidy, local_path_tidydata)
```


# Air dry moisture data
```{r}

name_of_transcription_file_on_drive = "weighing notes july 2,3,5 2026_transcribed20260806_AG" # copy the file's name exactly as it appears on google drive

dribble = drive_get(name_of_transcription_file_on_drive, shared_drive = 'Bogar Lab') # get necesary data for downloading

local_path <- paste0("Inputs/", 
                     name_of_transcription_file_on_drive, 
                     "_downloaded-", 
                     format(Sys.Date(), "%Y%m%d"), 
                     ".csv") #define what this local file name will be, with today's date

drive_download(dribble, type = "csv", path = local_path, overwrite = TRUE)

ad_transcribed_data = read_csv(local_path)

```



